ArKanjo: a tool for detecting function-level Code Duplication in the Linux Kernel

### 2025-07-03 paper/paper.tex
Minor adjustments

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The Linux kernel massive scale (+28 M LoC, +20 K contributors) presents unique 
maintenance challenges. Surprisingly, code duplication remains a persistent issue 
in the kernel codebase, which can hinder its evolution and patching. Academic 
approaches often focus on pairwise comparison of code artifacts, which are not directly applied 
to comprehensive codebase analyses. Other existing free
software\footnote{In this paper, we use the term \textit{Free Software}
interchangeably with \textit{Open Source Software}.} tools explored in 
practice frequently suffer from limited functionality, such as primitive textual 
matching, prove too narrow in scope, or fail to deliver effective results on complex, 
large-scale codebases. Existing solutions generally fail to address the Linux kernel 
specific needs: (1) scalability to handle its size, (2) actionable results for 
developers, and (3) integration with kernel development workflows.
%
This paper presents \textit{ArKanjo}, a novel command-line tool for Linux kernel maintenance 
designed to detect and analyze function-level duplications. Released under the MIT 
license, ArKanjo employs a two-stage architecture consisting of a Preprocessor and 
a Query Responder that separates computationally intensive analysis from efficient 
querying for duplications within large codebases. Pivotal advantages of ArKanjo over 
existing solutions include: (1) preprocessing that enables rapid queries without redundant analysis; 
and (2) prioritization of duplicates that impact maintainability, such as copied 
buggy logic. 
%
We evaluate ArKanjo against real-world duplication cases in recent 
kernel versions, demonstrating its effectiveness in identifying problematic clones 
that generic tools often overlook. By identifying well-defined, manageable 
duplication instances, ArKanjo effectively lowers the barrier for new contributors, 
a capability evidenced by its role in guiding students to make their first code 
improvements to the kernel. ArKanjo offers immediate value to kernel maintainers 
and serves as a replicable model for clone detection in other large-scale free 
software projects.

@@

The Linux kernel is a foundational free software project 
critical to a significant portion of the world digital infrastructure. Maintaining 
the kernel is an enormous undertaking involving over 28 million lines of code 
and contributions from over twenty thousand developers. Within this context, code 
duplication persists as a significant challenge, a known harmful practice that 
negatively affects code readability and increases the likelihood of introducing
bugs~\cite{harmone,harmtwo}. This issue is particularly acute in the 
kernel device drivers, which comprise over 66\% of the source code~\cite{marcelo}. 
For instance, maintainers of the AMD Display driver have specifically highlighted code 
duplication as a significant impediment to their work.

The detection of code duplication, or code clones, has been a research subject for 
decades~\cite{firstman}. The literature provides a standard taxonomy, classifying 
clones into four types based on their degree of similarity, from identical copies 
(Type-1) to semantically equivalent but textually different fragments (Type-4)~\cite{litreview}. 
Various detection methodologies have been developed, including textual, token-based, 
tree-based, and graph-based approaches~\cite{litreview}. These have culminated in 
state-of-the-art techniques, such as the graph-based work by Liu et al.~\cite{tailor}.

However, the primary focus of such academic work remains on determining whether a given 
pair of code artifacts are duplicates rather than providing a scalable method to scan 
an entire codebase for actionable results. Conversely, existing free software tools explored 
in practice, often lack this sophistication, relying instead on primitive textual matching 
that is insufficient for the complexity and scale of the kernel.

@@

We implemented a more straightforward approach for the difference method,
considering the number of exactly equal lines between two functions.
For two functions, \textit{function1} and \textit{function2}, we compute the similarity as the 
ratio of duplicated lines to the total number of lines across both functions.
This method is considerably simpler and more explainable than the text similarity method, 
The metric is defined by the proportion of common lines between the two functions. 
We use the \textit{diff} command built into the Linux environment to compute the number of equal lines.

@@

We propose three main functionalities for the user in the tool, called Duplication
Explorer, Function Information, and Duplication Report. Each functionality accepts
specific parameters to perform its operations.

@@

%Although a formal performance benchmark was not conducted, we recorded the tool's performance 
%on a machine equipped with a Ryzen 5700x processor and 32GB of RAM. The Preprocessor's 
%worst-case execution time across five runs was 8 minutes and 56 seconds on the AMD Display driver 
%and 9 minutes and 41 seconds on the IIO subsystem. In contrast, the Query Responder processed 
%requests in under two seconds, demonstrating the efficiency of the two-stage architecture 
%described in this paper.

While a formal performance benchmark was not conducted, we measured the Preprocessor's execution 
time on a machine equipped with a Ryzen 5700X processor and 32GB of RAM. Across five runs on the 
AMD Display driver, the worst-case execution time for the Preprocessor was 8 minutes and 56 seconds. 
On the IIO subsystem, the worst-case time over five runs was 9 minutes and 41 seconds. 
In contrast, the Query Responder processed requests in under two seconds in both cases, 
demonstrating the efficiency of the two-stage architecture.

@@

The student efforts demonstrated that newcomers could use the tool to contribute effectively. 
Of the 11 student patches, 5 were accepted, 5 were rejected, and 1 required follow-up 
fixes based on maintainer feedback. To remove the duplicated code, 7 student groups used the 
\textit{Parameterize Method}, and 2 used the \textit{Extract Method} \cite{refactorbook}, both of which are 
straightforward refactoring strategies. These results support the conclusion that ArKanjo identifies 
actionable duplications and that newcomers can successfully contribute patches to the kernel.

While this research includes successfully merged patches by the main author and student groups,
a significant portion of the student contributions encountered notable challenges. These difficulties 
often stemmed from maintainer feedback indicating that, although the proposed changes reduced duplication, 
they were rejected or required substantial revision due to concerns about code readability, increased abstraction, 
or the specific context of the code. Additional challenges included technical complexities, such as C macros in 
configuration files and extended patch review timelines. These findings highlight that while ArKanjo 
effectively identifies code duplications, a successful 
contribution also requires navigating trade-offs between deduplication and other factors, such as 
contextual appropriateness and maintainer expectations.

@@

This paper presented ArKanjo, a novel tool designed to detect effectively 
and facilitate the refactoring of function-level code duplication within the Linux kernel. 
Evaluations demonstrated that ArKanjo is highly effective at identifying Type-1 and Type-2 
clones, which are often indicative of propagated issues such as copied buggy logic. We validated the practical 
value of the tool through an ethnographic study in which the main author and student 
groups used ArKanjo to identify actionable duplications. This effort led to 13 patch submissions, 
including one from the main author and five from student groups that were successfully merged into 
the kernel
\footnote{Reference for the accepted patches:  
\textit{\href{https://github.com/arkanjo-tool/arkanjo/blob/main/PATCHES.md}{github.com/arkanjo-tool/arkanjo/blob/main/PATCHES.md}}.
}, 
collectively removing 991 lines of duplicated code. While some refactoring 
efforts were declined due to maintainer priorities, ArKanjo successfully lowered the barrier 
for new contributors to make meaningful improvements. The tool provides immediate value for 
kernel maintenance and serves as a model for clone detection in other large-scale projects.

@@

This study was financed, in part, by CAPES (Finance Code 001), the University of São Paulo – USP (Proc. 22.1.9345.1.2), the São Paulo Research Foundation – FAPESP and the São Paulo State Data Analysis System Foundation – SEADE (Proc. 2023/18026-8), Brazil.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The Linux kernel massive scale (+28 M LoC, +20 K contributors) presents unique 
maintenance challenges. Surprisingly, code duplication remains a persistent issue 
in the kernel codebase, which can hinder its evolution and patching. Academic 
approaches often focus on pairwise comparison of code artifacts, which are not directly applied 
to comprehensive codebase analyses. Other existing free
software\footnote{In this paper, we use the term \textit{Free Software}
interchangeably with \textit{Open Source Software}.} tools explored in 
practice frequently suffer from limited functionality, such as primitive textual 
matching, prove too narrow in scope, or fail to deliver effective results on complex, 
large-scale codebases. Existing solutions generally fail to address the Linux kernel-specific needs: (1) scalability to handle its size, (2) actionable results for 
developers, and (3) integration with kernel development workflows.
%
This paper presents \textit{ArKanjo}, a novel command-line tool for Linux kernel maintenance 
designed to detect and analyze function-level duplications. Released under the MIT 
license, ArKanjo employs a two-stage architecture consisting of a Preprocessor and 
a Query Responder that separates computationally intensive analysis from efficient 
querying for duplications within large codebases. Pivotal advantages of ArKanjo over 
existing solutions include: (1) preprocessing that enables rapid queries without redundant analysis; 
and (2) prioritization of duplicates that impact maintainability, such as copied 
buggy logic. 
%
We evaluate ArKanjo against real-world duplication cases in recent 
kernel versions, demonstrating its effectiveness in identifying problematic clones 
that generic tools often overlook. By identifying well-defined, manageable 
duplication instances, ArKanjo effectively lowers the barrier for new contributors, 
a capability evidenced by its role in guiding students to make their first code 
improvements to the kernel. ArKanjo offers immediate value to kernel maintainers 
and serves as a replicable model for clone detection in other large-scale free 
software projects.

@@

The Linux kernel is a foundational free software project 
critical to a significant portion of the world digital infrastructure. Maintaining 
the kernel is an enormous undertaking involving over 28 million lines of code 
and contributions from over twenty thousand developers. Within this context, code 
duplication persists as a significant challenge, a known harmful practice that 
negatively affects code readability and increases the likelihood of introducing
bugs~\cite{harmone,harmtwo}. This issue is particularly acute in kernel device drivers, which comprise over 66\% of the source code~\cite{marcelo}. 
For instance, maintainers of the AMD Display driver have specifically highlighted code duplication as a significant impediment to their work.

Detecting code duplication, or code clones, has been a research subject for 
decades~\cite{firstman}. The literature provides a standard taxonomy, classifying 
clones into four types based on their degree of similarity, from identical copies 
(Type-1) to semantically equivalent but textually different fragments (Type-4)~\cite{litreview}. 
Various detection methodologies have been developed, including textual, token-based, 
tree-based, and graph-based approaches~\cite{litreview}. These have culminated in 
state-of-the-art techniques, such as the graph-based work by Liu et al.~\cite{tailor}.

However, the primary focus of such academic work remains on determining whether a given pair of code artifacts is a duplicate rather than providing a scalable method to scan 
an entire codebase for actionable results. Conversely, existing free software tools explored 
in practice, often lack this sophistication, relying instead on primitive textual matching 
that is insufficient for the complexity and scale of the kernel.

@@

We implemented a more straightforward approach for the difference method,
considering the number of exactly equal lines between two functions.
For two functions, \textit{function1} and \textit{function2}, we compute the similarity as the 
ratio of duplicated lines to the total number of lines across both functions.
This method is considerably simpler and more explainable than the text similarity method. 
The metric is defined by the proportion of common lines between the two functions. 
We use the \textit{diff} command built into the Linux environment to compute the number of equal lines.

@@

We propose three main functionalities for the user in the tool: Duplication Explorer, Function Information, and Duplication Report. Each functionality accepts
specific parameters to perform its operations.

@@

While we did not conduct a formal performance benchmark, we measured the Preprocessor's execution 
time on a machine equipped with a Ryzen 5700X processor and 32GB of RAM. Across 5 runs on the 
AMD Display driver, the worst-case execution time for the Preprocessor was 8 minutes and 56 seconds. 
On the IIO subsystem, the worst-case time over 5 runs was 9 minutes and 41 seconds. 
In contrast, the Query Responder processed requests in under 2 seconds in both cases, 
demonstrating the efficiency of the two-stage architecture.

@@

The students' efforts demonstrated that newcomers could use the tool to contribute effectively. 
Of the 11 students' patches, 5 were accepted, 5 were rejected, and 1 required follow-up 
fixes based on maintainer feedback. To remove the duplicated code, 7 student groups used the 
\textit{Parameterize Method}, and 2 used the \textit{Extract Method} \cite{refactorbook}, both of which are 
straightforward refactoring strategies. These results support that ArKanjo identifies 
actionable duplications and that newcomers can successfully contribute patches to the kernel.

While this research includes successfully merged patches by the main author and student groups,
a significant portion of the students' contributions encountered notable challenges. These difficulties 
often stemmed from maintainer feedback indicating that, although the proposed changes reduced duplication, 
they were rejected or required substantial revision due to concerns about code readability, increased abstraction, 
or the specific context of the code. Additional challenges included technical complexities, such as C macros in 
configuration files and extended patch review timelines. These findings highlight that while ArKanjo 
effectively identifies code duplications, a successful 
contribution also requires navigating trade-offs between deduplication and other factors, such as 
contextual appropriateness and maintainer expectations.

@@

This paper presented ArKanjo, a novel tool designed to detect effectively and facilitate the refactoring of function-level code duplication within the Linux kernel. 
Evaluations demonstrated that ArKanjo is highly effective at identifying Type-1 and Type-2 
clones, which are often indicative of propagated issues such as copied buggy logic. We validated the practical 
value of the tool through an ethnographic study in which the main author and student groups used ArKanjo to identify actionable duplications. This effort led to 13 patch submissions, including 1 from the main author and 5 from student groups that were successfully merged into the kernel
\footnote{Reference for the accepted patches:  
\textit{\href{https://github.com/arkanjo-tool/arkanjo/blob/main/PATCHES.md}{github.com/arkanjo-tool/arkanjo/blob/main/PATCHES.md}}.
}, collectively removing 991 lines of duplicated code. While some refactoring efforts were declined due to maintainer priorities, ArKanjo successfully lowered the barrier 
for new contributors to make meaningful improvements. The tool provides immediate value for 
kernel maintenance and serves as a model for clone detection in other large-scale projects.

@@

This study was financed, in part, by CAPES (Finance Code 001), the University of São Paulo – USP (Proc. 22.1.9345.1.2) and the São Paulo Research Foundation – FAPESP with the São Paulo State Data Analysis System Foundation – SEADE (Proc. 2023/18026-8), Brazil.
